Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow

### 2009-05-15 paper/paper.tex
More hacking on paper.

\begin{abstract}
Forensic analysis requires the acquisition and management of many
different types of evidence, including individual disk drives, RAID
sets, network packets, memory images, and extracted files. Often the
same evidence is reviewed by several different tools or examiners in
different locations. We propose a backwards-compatible redesign of the
Advanced Forensic Format---an open, extensible file format for storing
and sharing of evidence, arbitrary case related information and
analysis results among different tools. The new specification, termed
AFF4, was designed to be simple to implement, bult upon the well
supported ZIP file format specification. AFF4 implements will have
downward compatability with existing AFF files.
\end{abstract}

@@

Both the AFF and EWF file formats are designed to store a single
image, and any metadata implicitely refers to that image. Therefore,
these formats in themselves have a limited metadata model which is
unable to exist outside the context of the image itself.
But AFF went beyond merely
specifying metadata as an addition to the forensic evidence
itself. Metadata forms an essential part of the file format, with the
same metadata system storing both essential information required to
read the image itself, as well as non-essential metadata. AFF allows
the use of arbitrary name/value pairs to store metadata, but it also
stored critical information such as sector size and device serial
number using the same metadata representation system. \verb+Aimage+,
the AFF hard  disk acquisition tool should, not only stored the image,
but also stored a description of the tool
itself, the version of AFFLIB used to create the image, the computer
on which the image was made, the operator of the tool, the user supplied parameters
supplied to the tool, all in addition to the drive being imaged.

@@

We call the new system AFF4, and use the phrase AFF1 to refer to the
legacy system developed by Garfinkel et al.\footnote{Although Garfinkel
never changed the AFF bit-level specification, Garfinkel released
AFFLIB implementations with major version numbers 1, 2 and 3. We
therefore call our system AFF4 to avoid confusion.} The AFF4
implementation, when it is publicly released, will be able to read
existing AFF files.

@@

\section{Global Distributed Evidence Management}
While AFF1 was designed for use on a single machine that could both
image evidence and perform analysis, many modern practitioners work in
distributed environments in which imaging and analysis takes place in
multiple locations and is performed by multiple individuals.

@@

\subsection{Migration from AFF1 to AFF4}
We have addressed these issues in a number of ways. First, we have
extended the AFF schema to support multiple forensic images (which are
now called \emph{streams}). It is therefore possible now to store
multiple related streams in the same volume, or to acquire mutiple
streams at the same time (e.g. network and disk data).

We replaced AFF's ``bit-level'' format with an improved format that is
more flexible and has better existing support, namely the Zip64 file
format is now adopted as a default storage format. This enhanced our
compatibility with other tools and simplifies implementations for
compatible systems.

We have also defined an information model which addresses uniquely
identifying evidence data, expressing arbitrary information and
integrating information and image data from multiple evidence volumes
at a time. This allows us to extend evidence collection and management
to a truely universal scale---allowing creation and management of a
large corpus.

We have also released a library and toolset, \emph{LIBAFF4}
implementing the new schema and file format. This new library provides
legacy support for existing AFF1 volumes.

@@

\item \emph{A Relation} is a statement about an AFF Object which
expresses some property of the object. The relation comprises of a
triple of (Subject, Attribute, Value). All metadata is reduced to the
triple notation.

@@

Management of evidence requires an effective identififaction, with
practitioners currently employing acquisition time metadata such as
case identifiers and description fields in the EWF file format; file
and directory naming schemes, and labeling of evidence container hard
drives. Evidence may also be referenced by external means in an
inconsistent way. For example, in an investigator's case note a disk
image may be referred to by the name of the suspect (e.g. Joe's hard
disk), the case number or dates.

@@

The AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. An \emph{Attribute} can be any kind of value, such
as the sector size of a device, a device capacity, or the name of the
person who performed an imaging operation. A \emph{Value} is the value
of the attribute. Using this system we are able to store
arbitrary attributes about any object in the AFF4
universe. Additionally, as these statements are universally scoped,
they may be stored anywhere.

The AFF4 design extends beyond the management of a single volume,
stream or image to a universal system for managing data of many
types. This necessarily means that a single running instance is
generally unable to have visibility of the entire AFF4 universe. For
example, if a volume is opened which contains a Map Stream targeting a
stream stored in a different volume, it is not generally possible to
tell in which volume the stream exists, and where that volume is
actually stored.

@@

To this end we define a way for serializing metadata statements (or
triples) into a standard format which implementations can load into
their respective resolvers when parsing the volume. Triples can be
stored in segments having a URN ending with ``{\em properties}''. The
AFF4 implementation loads these segments automatically into the
Universal Resolver.

Triples are stored within the properties segment one per line, with
the subject URN (encoded according to RFC1737), followed by whitespace
and the attribute name. This is then followed by the equal sign and
the UTF8 encoding of the value. An example properties file for an
Image Stream is shown in Listing \ref{image_metadata}.

@@

When the volume is loaded, the AFF4 implementation automatically loads
any properties files and populates its Universal Resolver with the
information visibile to it. AFF4 provides a mechanism to use an
external resolver as well---for example, we have implemented a
resolved that stores Attributes in a MySQL database to provide for a
Universal Resolver that shares information between different instances
on the same network.

@@

Link Object:
  map aff4:target=urn:aff4:ed8f1e7a-94aa
  map aff4:type=link
\end{lstlisting}

@@

\subsection{Unused Zip64 Features}
There are a number of aspects in the Zip file format specifications
which we do not utilize.
\begin{itemize}
\item We ignore Zip64's built-in support for splitting
archives into multiple Zip files. Instead, our implementation treats
each volume as a complete and stand-alone Zip file. The AFF4
implementation then considers the segments contained within as
belonging to the universal collection. We provide the ability to split
a stream across volumes automatically, as each segment may be stored
in a different volume, and the Universal Resolver is used to find it.

\item Although Zip64 also defines encryption and authentication
extensions, we do not use them due to the restrictions imposed on
their use and because they lack the functionality that is important
for a forensic user. Instead, we use AFF4's digital signature
facilities for integrity and non-repudiation, and we introduce a new
stream based encryption scheme for ensuring data privacy
(Section~\ref{crypted_stream}).

\subsection{Implementing Zip64 in AFFLIB}
Although there are numerous Zip implementations available today, we
have created our own implementation. There are many reasons to develop
our own Zip64 implementation for AFF4:

@@

For this reason, URLs may be used interchangeably with a URN within
the AFF4 universe. For example, the \emph{aff4:stored} attribute of a
volume may be specified as a URL
(e.g. \emph{http://intranet/123453/}).  AFFLIB provides transparent
support for HTTP and FTP URLs by means of the curl HTTP
library\cite{libcurl}. The HTTP Stream, therefore satisfies read
requests by making HTTP requests to the web server. We use the
\emph{Content-Range} HTTP header to request exactly the byte range the
client is interested in. This allows efficient network transport as we
do not need to download unnecessary data, we just request exactly the
data the client application requires.

Our implementation also enables direct writing to a HTTP URL using the
WebDav extensions to HTTP\cite{webdav-rfc}. The HTTP stream also
supports the File Transfer Protocol (FTP) and HTTPS (Secure Sockets
Layer---SSL) protocols transparently. This support is provided
automatically by the curl library.

@@

\subsection{Identity Object}
AFF4 defines a \emph{Statement} as a collection of facts, or
(subject, attribute, value) tuples. Listing \ref{statement}
illustrates a collection of facts encoded in the standard AFF4
notation (SHA254 hashes are base64 encoded).

\begin{lstlisting}[
	caption=An Example Statement  (URNs and hashes are
shortened for illustration).,
	label=statement,
	float=tb,
	 ]
urn:aff4:34a62f06/00000000 aff4:sha256=+Xf4i....7rPCgo=
urn:aff4:34a62f06/00000000.idx aff4:sha256=ptV7xOK6....C7R6Xs=
urn:aff4:34a62f06/properties aff4:sha256=yoZ9dtv....eUignYMtk=
urn:aff4:34a62f06 aff4:sha256=udajC5C...BVii7psU=
\end{lstlisting}

@@

\subsection{Private Streams}
Often forensic software needs to store internal state, such as the
current state of the GUI, or internal data structures which may be
required for caching.

Currently, software create additional cache files to maintain this
information. It is advantageous to be able to store these in the
evidence volume itself. The evidence volume can then be re-opened by
the software at a later stage and private data can be retrieved and
used directly. Such private application data can be stored in a
Private stream. There is no specification of what can be stored in the
private stream, and applications with private streams should be able
to store arbitrary segments in any format at all. Segments URNs should
be prefixed by the private stream URNs, and the stream type should be
specified as ``private''.

It is recommended that applications create attributes for the private
stream so they can reliably recognize their own private stream, and
identify the version of the application which created it. The AFFLIB
API can be used to render access to private segments and private
attributes.

>>>>

\begin{abstract}
Forensic analysis requires the acquisition and management of many
different types of evidence, including individual disk drives, RAID
sets, network packets, memory images, and extracted files. Often the
same evidence is reviewed by several different tools or examiners in
different locations. We propose a backwards-compatible redesign of the
Advanced Forensic Format---an open, extensible file format for storing
and sharing of evidence, arbitrary case related information and
analysis results among different tools. The new specification, termed
AFF4, was designed to be simple to implement, bult upon the well
supported ZIP file format specification. AFF4 implementation have
downward compatability with existing AFF files.
\end{abstract}

@@

Both the AFF and EWF file formats are designed to store a single
image, and any metadata implicitely refers to that image. Therefore,
these formats in themselves have a limited metadata model which is
unable to exist outside the context of the image itself.  But AFF went
beyond merely specifying metadata as an addition to the forensic
evidence itself. Metadata forms an essential part of the file format,
with the same metadata system storing both essential information
required to read the image itself, as well as non-essential
metadata. AFF allows the use of arbitrary name/value pairs to store
metadata, but it also stored critical information such as sector size
and device serial number using the same metadata representation
system. \verb+Aimage+, the AFF hard disk acquisition tool, not only
stores the image, but additionally stores a description of the tool
itself, the version of AFFLIB used to create the image, the computer
on which the image was made, the operator of the tool, the user
supplied parameters supplied to the tool.

@@

We call the new system AFF4, and use the phrase AFF1 to refer to the
legacy system developed by Garfinkel et al.\footnote{Although
Garfinkel never changed the AFF bit-level specification, Garfinkel
released AFFLIB implementations with major version numbers 1, 2 and
3. We therefore call our system AFF4 to avoid confusion.} The publicly
released AFF4 implementation, is able to read existing AFF files.

@@

\subsection{Global Distributed Evidence Management}
While AFF1 was designed for use on a single machine that could both
image evidence and perform analysis, many modern practitioners work in
distributed environments in which imaging and analysis takes place in
multiple locations and is performed by multiple individuals.

@@

%\subsection{Migration from AFF1 to AFF4}
%We have addressed these issues in a number of ways. First, we have
%extended the AFF schema to support multiple forensic images (which are
%now called \emph{streams}). It is therefore possible now to store
%multiple related streams in the same volume, or to acquire mutiple
%streams at the same time (e.g. network and disk data).

%We replaced AFF's ``bit-level'' format with an improved format that is
%more flexible and has better existing support, namely the Zip64 file
%format is now adopted as a default storage format. This enhanced our
%compatibility with other tools and simplifies implementations for
%compatible systems.

%We have also defined an information model which addresses uniquely
%identifying evidence data, expressing arbitrary information and
%integrating information and image data from multiple evidence volumes
%at a time. This allows us to extend evidence collection and management
%to a truely universal scale - allowing creation and management of a
%large corpus.

%We have also released a library and toolset, \emph{LIBAFF4}
%implementing the new schema and file format. This new library provides
%legacy support for existing AFF1 volumes.

@@

\item \emph{A Relation or Fact} is a statement about an AFF Object which
  expresses some property of the object. The relation comprises of a
  tuple of (Subject, Attribute, Value). All metadata is reduced to
  this tuple notation.

@@

Management of evidence requires an effective identification, with
practitioners currently employing acquisition time metadata such as
case identifiers and description fields in the EWF file format; file
and directory naming schemes, and labeling of evidence container hard
drives. Evidence may also be referenced by external means in an
inconsistent way. For example, in an investigator's case note a disk
image may be referred to by the name of the suspect (e.g. Joe's hard
disk), the case number or dates.

@@

The AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in tuple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. An \emph{Attribute} can be any kind of value, such
as the sector size of a device, a device capacity, or the name of the
person who performed an imaging operation. A \emph{Value} is the value
of the attribute. Using this system we are able to store
arbitrary attributes about any object in the AFF4
universe. Additionally, as these statements are universally scoped,
they may be stored anywhere.

The AFF4 design extends beyond the management of a single volume,
stream or image to a universal system for managing data of many
types. This necessarily means that a single running instance is
generally unable to have visibility of the entire AFF4 universe. For
example, if a volume is opened which contains a Map Stream targeting a
stream stored in a different volume, it is not generally possible to
tell where that volume is actually stored.

@@

To this end we define a way for serializing metadata statements (or
tuples) into a standard format which implementations can load into
their respective resolvers when parsing the volume. Relations can be
stored in segments having a URN ending with ``{\em properties}''. The
AFF4 implementation loads these segments automatically into the
Universal Resolver.

Relations are stored within the properties segment one per line, with
the subject URN (encoded according to RFC1737), followed by whitespace
and the attribute name. This is then followed by the equal sign and
the UTF8 encoding of the value. An example properties file for an
Image Stream is shown in Listing \ref{image_metadata}.

@@

When the volume is loaded, the AFF4 implementation automatically loads
any properties files and populates its Universal Resolver with the
information visibile to it. AFF4 provides a mechanism to use an
external resolver as well---for example, we have implemented a
resolver that stores Attributes in a MySQL database to provide for a
Universal Resolver that shares information between different instances
on the same network.

@@

Link Object:
  map aff4:target=urn:aff4:ed8f1e7a-94aa
  map aff4:type=link

Identity Object:
  urn:aff4:identity/41:13 aff4:common_name=/C=US/ST=CA/
                          L=SanFrancisco/O=Fort-Funston/
			  CN=client1/emailAddress=
			  me@myhost.mydomain
  urn:aff4:identity/41:13 aff4:type=identity
  urn:aff4:identity/41:13 aff4:statement=00000000
  urn:aff4:identity/41:13 aff4:x509=urn:aff4:identity/
                          41:13/cert.pem
\end{lstlisting}

@@

We ignore Zip64's built-in support for splitting archives into
multiple Zip files. Instead, our implementation treats each volume as
a complete and stand-alone Zip file. The AFF4 implementation then
considers the segments contained within as belonging to the universal
collection. This provide the ability to split a stream across volumes
automatically, as different segments within the same stream may be
stored in different volumes.

Zip64 also defines encryption and authentication extensions. We do not
use them due to the restrictions imposed on their use and because they
lack the functionality that is important for a forensic user. Instead,
we use AFF4's digital signature facilities for integrity and
non-repudiation, and we introduce a new stream based encryption scheme
for ensuring data privacy (Section~\ref{crypted_stream}).

Although there are numerous Zip implementations available today, we
have created our own implementation. There are many reasons to develop
our own Zip64 implementation for AFF4:

@@

For this reason, URLs may be used interchangeably with a URN within
the AFF4 universe. For example, the \emph{aff4:stored} attribute of a
volume may be specified as a URL
(e.g. \emph{http://intranet/123453/}).  AFF4 provides transparent
support for HTTP and FTP URLs by means of the curl HTTP
library\cite{libcurl}. The HTTP Stream, therefore satisfies read
requests by making HTTP requests to the web server. We use the
\emph{Content-Range} HTTP header to request exactly the byte range the
client is interested in. This allows efficient network transport as we
do not need to download unnecessary data, we just request those chunks
the client application requires.

Our implementation also enables direct writing to a HTTP URL using the
WebDav extensions to HTTP\cite{webdav-rfc}. The HTTP stream also
supports the File Transfer Protocol (FTP) and HTTPS (Secure Sockets
Layer---SSL) protocols transparently, as provided by the curl library.

@@

\section{Identity Object}
AFF4 defines a \emph{Statement} as a collection of facts, or
(subject, attribute, value) tuples. Listing \ref{statement}
illustrates a collection of facts encoded in the standard AFF4
notation (SHA254 hashes are base64 encoded).

\begin{lstlisting}[
	caption=An Example Statement  (URNs and hashes are
shortened for illustration).,
	label=statement,
	float=tb,
	 ]
urn:aff4:34a62f06/00000 aff4:sha256=+Xf4i....7rPCgo=
urn:aff4:34a62f06/00000.idx aff4:sha256=ptV7xOK6....C7R6Xs=
urn:aff4:34a62f06/properties aff4:sha256=yoZ....YMtk=
urn:aff4:34a62f06 aff4:sha256=udajC5C...BVii7psU=
\end{lstlisting}

@@

%\section{Private Objects}
%Often forensic software needs to store internal state, such as the
%current state of the GUI, or internal data structures which may be
%required for caching.

%Currently, software create additional cache files to maintain this
%information. It is advantageous to be able to store these in the
%evidence volume itself. The evidence volume can then be re-opened by
%the software at a later stage and private data can be retrieved and
%used directly. Such private application data can be stored in a
%Private object. There is no specification of what can be stored in the
%private object, and applications with private objects should be able
%to store arbitrary segments in any format at all. Segments URNs should
%be prefixed by the private object's URN, and the stream type should be
%specified as ``private''.

%It is recommended that applications create attributes for the private
%stream so they can reliably recognize their own private stream, and
%identify the version of the application which created it. The AFF4
%API can be used to render access to private segments and private
%attributes.
